---
output:
  word_document: default
  html_document: default
---
# MIS503 - Final Project
## Pierce, Michaela
### Zillow Home Value Index Analysis


```{r}
library(readr)
library(tidyr)
library(dplyr)
library(ggplot2)
```




### Wake County Home Sales

**a. What have been the overall trends in Wake County Home Values?**
The home values have increased over the years. 

**b. There were dips in home values in the past 10 years. What years did these occur?**  
2013 - 2021



**c. Based on the analysis, where would be the least expensive area to purchase home? Most expensive area?**  
Least Expensive: Zebulon ; Most expensive: Cary


**d. What has happened to the overall property values in Apex and Cary in 2023?**  
Apex and Cary's overall property values did not increase in 2023. 


```{r}
# load data
Sales <- read_csv("SingleFamilyResidenceSales.csv")

# Create the WakeCountySales tibble
WakeCountySales <- Sales %>%
  filter(State == "NC" & CountyName == "Wake County") %>%
  select(RegionName, State, CountyName, Metro,
         `5/31/2013` = `5/31/2013`,
         `5/31/2014` = `5/31/2014`,
         `5/31/2015` = `5/31/2015`,
         `5/31/2016` = `5/31/2016`,
         `5/31/2017` = `5/31/2017`,
         `5/31/2018` = `5/31/2018`,
         `5/31/2019` = `5/31/2019`,
         `5/31/2020` = `5/31/2020`,
         `5/31/2021` = `5/31/2021`,
         `5/31/2022` = `5/31/2022`,
         `5/31/2023` = `5/31/2023`)

# Rename date columns to only include year
colnames(WakeCountySales)[5:16] <- gsub("5/31/", "", colnames(WakeCountySales)[5:16])

# View the tibble
# View(WakeCountySales)

```

```{r}


# Tidy the data
TidyWakeCountySales <- WakeCountySales %>%
  pivot_longer(cols = matches("^\\d{4}$"), 
               names_to = "YR",             
               values_to = "ZHVI")         

# View the data
# View(TidyWakeCountySales)



```


```{r}
# Convert YR to numeric
TidyWakeCountySales$YR <- as.numeric(as.character(TidyWakeCountySales$YR))

# Create scatter plot
scatter_plot <- ggplot(TidyWakeCountySales, aes(x = YR, y = ZHVI, color = RegionName)) +
  geom_point() +
  scale_x_continuous(name = "Year", breaks = unique(TidyWakeCountySales$YR)) +  # Show every year
  scale_y_continuous(name = "Home Values (in $)", labels = scales::comma) +  # Add commas to y-axis labels
  labs(title = "Wake County Home Sales",
       x = "Year",
       y = "Home Value (ZHVI)") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))  

# Display the scatter plot
print(scatter_plot)


```








```{r}

# Create stacked bar graph
stacked_bar <- ggplot(TidyWakeCountySales, aes(x = factor(YR), y = ZHVI, color = RegionName)) +
  geom_col() +
  scale_x_discrete(name = "Year", labels = as.character(unique(TidyWakeCountySales$YR))) +  # show all years on x-axis
  scale_y_continuous(name = "Home Values (in $)", labels = scales::comma) +  # Add commas to y-axis labels
  labs(title = "Wake County Home Sales",
       y = "Total Home Value (ZHVI)") +  # Corrected the y-axis label
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))
 

# Display stacked bar graph
print(stacked_bar)

```



---


### NC Rental Market 

**a. What has been the overall trend in the rental market around the state? Are there any cities that have not followed this trend?**  
The rental market has increased over the years. Fayetteville remained consistent until 2021 where it spiked up. 

**b. Where is the most expensive city to rent in? Least expensive?**   
Most Expensive: Asheville ; Least Expensive: Fayetteville


**c. You are trying to decide between Wilmington and Asheville. Which market has the lowest rent?**  
Wilmington has the lowest rent.


```{r}
# load data
Rentals <- read_csv("SingleFamilyResidenceRental.csv")

# Asheville, Charlotte, Durham, Fayetteville, Raleigh, and Wilmington
NC_RentalCities <- c("Asheville", "Charlotte", "Durham", "Fayetteville", "Raleigh", "Wilmington")


# Filter data
Rentals <- Rentals %>%
  filter(RegionName %in% NC_RentalCities,
         State == "NC") %>%
  select(RegionName, State, matches("^1/31/20\\d{2}$"))

# Rename date columns to only include the year
colnames(Rentals)[3:10] <- gsub("1/31/", "", colnames(Rentals)[3:10])

# View the tibble
# View(Rentals)


```
```{r}

# Tidy the data
Tidy_Rentals <- Rentals %>%
  pivot_longer(cols = matches("^20\\d{2}$"),
               names_to = "YR",
               values_to = "ZHVI")

# View the tibble
# View(Tidy_Rentals)
```

```{r}

# Create scatter plot
scatter_plot <- ggplot(Tidy_Rentals, aes(x = as.numeric(YR), y = ZHVI, color = RegionName)) +
  geom_point() +
  scale_x_continuous(name = "Year", breaks = unique(as.numeric(Tidy_Rentals$YR))) +  
  scale_y_continuous(name = "Home Value (ZHVI)", labels = scales::comma) +
  labs(title = "Home Values by Year and City",
       x = "Year",
       y = "Average Rentals Cost") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5))  

# Display the scatter plot
print(scatter_plot)

```


```{r}
# Create stacked bar graph
stacked_bar <- ggplot(Tidy_Rentals, aes(x = as.numeric(YR), y = ZHVI, color = RegionName)) +
  geom_col() +
  scale_x_continuous(name = "Year", breaks = unique(as.numeric(Tidy_Rentals$YR))) +  
  scale_y_continuous(name = "Average Rental Cost", labels = scales::comma) +
  labs(title = "Rental Costs in NC Cities",
       x = "Year",
       y = "Average Rental Cost") +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5)) 

# Display the stacked bar graph
print(stacked_bar)

```






---



### Home Values in Select Markets

**a. According to the results, which market has the lowest median price (represented as horizontal bar in box plot)?**   
Charlotte-Concord-Gastonia



**b. The violin plot will show density meaning the wider the plot is, the more observations occur within that area. Which market has the most density around the median value of homes?**   
Wilmington



**c. The box plot will also show outliers in the various markets. Which metro area had the largest outlier (i.e., the highest value home sold in the past 10 years)?**   
Raleigh-Cary and Asheville


```{r}
#  Asheville, Charlotte-Concord-Gastonia, Raleigh-Cary and Wilmington
NCHomeSales_Cities <- c("Asheville", "Charlotte-Concord-Gastonia", "Raleigh-Cary", "Wilmington")

# Create the NCHomeSales tibble
NCHomeSales <- Sales %>%
  filter(State == "NC" & Metro %in% NCHomeSales_Cities) %>%
  select(RegionName, State, Metro,
         `5/31/2013` = `5/31/2013`,
         `5/31/2014` = `5/31/2014`,
         `5/31/2015` = `5/31/2015`,
         `5/31/2016` = `5/31/2016`,
         `5/31/2017` = `5/31/2017`,
         `5/31/2018` = `5/31/2018`,
         `5/31/2019` = `5/31/2019`,
         `5/31/2020` = `5/31/2020`,
         `5/31/2021` = `5/31/2021`,
         `5/31/2022` = `5/31/2022`,
         `5/31/2023` = `5/31/2023`)

# Rename date columns to only include year
colnames(NCHomeSales)[4:ncol(NCHomeSales)] <- gsub("5/31/", "", colnames(NCHomeSales)[4:ncol(NCHomeSales)])

# View the tibble
# View(NCHomeSales)


```


```{r}

# Tidy the data.
TidyNCHomeSales2 <- NCHomeSales %>%
  gather(key = "Year", value = "ZHVI", -RegionName, -State, -Metro) %>%
  mutate(Year = gsub("\\.", "/", Year)) 

# Group by Metro
NCHomeSales_grouped <- TidyNCHomeSales2 %>%
  group_by(Metro)

```



```{r}

ggplot(NCHomeSales_grouped, aes(x = Metro, y = ZHVI)) +
  geom_violin(width = 0.7) +  # Adjust the width of the violin plot
  geom_boxplot(width = 0.2) +  # Adjust the width of the box plot
  scale_y_continuous(breaks = c(0, 1000000, 2000000), labels = c("0", "1M", "2M")) +
  labs(x = "Metro", y = "Home Values (in $)") +
  theme(axis.text.x = element_text(angle = 0)) 



```





---


#  Relocation Home Value Comparison



**a. Based on your analysis, which city’s housing is most affordable? Least affordable?**   
Most affordable city: Houston ; Least affordable city: New York

**b. Which cities saw the largest change in prices over the past 5 years? Which city has remained more consistent (i.e., no huge swings up or down in home values)?**   
City with largest change in prices over the past 5 years: Denver  ;   City that has remained consistent over the past 5 years: Chicago

**c. Which cities saw a decline in value during 2023 and which cities remained consistent?**   
Denver saw a decline in home values during 2023.New York had a very small decline. ; Chicago and Houston's home values remained consistent during 2023.



```{r}
# Filter data
NationalHomeSales <- Sales %>%
  filter(RegionName %in% c("Chicago", "Denver", "Houston", "New York"),
         CountyName %in% c("Cook County", "Denver County", "Harris County", "Queens County")) %>%
  select(RegionName, State, CountyName, Metro,
         `5/31/2013`, `5/31/2014`, 
         `5/31/2015`, `5/31/2016`, 
         `5/31/2017`, `5/31/2018`, 
         `5/31/2019`, `5/31/2020`, 
         `5/31/2021`, `5/31/2022`, 
         `5/31/2023`)

# Rename the date columns to only include the year
colnames(NationalHomeSales)[5:ncol(NationalHomeSales)] <- gsub("5/31/", "", colnames(NationalHomeSales)[5:ncol(NationalHomeSales)])

# View the tibble
# View(NationalHomeSales)

```


```{r}
# Tidy the data
NationalHomeSales_tidy <- NationalHomeSales %>%
  pivot_longer(cols = `2013`:`2023`, names_to = "YR", values_to = "ZHVI")

# View the data
# View(NationalHomeSales_tidy)
```


```{r}
library(scales)

# Plotting scatterplots for each city separately using facets
ggplot(NationalHomeSales_tidy, aes(x = YR, y = ZHVI)) +
  geom_point() +
  facet_wrap(~Metro) +
  labs(x = "YR", y = "Home Values (in $)") +
  scale_y_continuous(labels = label_comma()) +  # Add commas to y-axis labels
  theme(axis.text.x = element_text(angle = 90, hjust = 1))  

```



---


### Future Home Values

**a. Which is the only city that is projected to have a decrease in home values in the next 3 months?**
No city projected to have a decrease in home values in the next 3 months


**b. If you are only concerned about the largest home value increase (by percentage) in the next 12 months, which city would you choose to relocate to?**
Chicago, IL



```{r}
Projections <- read_csv("Projections.csv")

# Filter the data for the specified cities and columns
FutureHomeValues <- Projections %>%
  filter(RegionName %in% c("Chicago, IL", "Denver, CO", "Houston, TX", "New York, NY")) %>%
  select(RegionName, `2024-04-30`, `2024-06-30`, `2025-03-31`)

# Rename the columns
colnames(FutureHomeValues)[2:4] <- c("Current", "ThreeMonths", "TwelveMonths")

# View the tibble
# View(FutureHomeValues)



```

```{r}

# Tidy the data
FutureHomeValues_tidy <- FutureHomeValues %>%
  pivot_longer(cols = -RegionName, names_to = "Time", values_to = "PercentageChange")

# Plotting scatter plot with city as color
ggplot(FutureHomeValues_tidy, aes(x = Time, y = PercentageChange, color = RegionName)) +
  geom_point() +
  labs(x = "Time", y = "Percentage Change") +
  theme(axis.text.x = element_text(angle = 0))
```











