---
output:
  word_document: default
  html_document: default
---
# Module 6 - Assignment 1
## Pierce, Michaela 
### Lubridate




```{r}
# Load necessary packages
library(tidyverse)
library(lubridate)

```




```{r}
library(readr)
Appointments <- read_csv("Appointments.csv")
#View(Appointments)
```



# Working with Dates


```{r}
# set the ReservationYear to 2019 for all the rows
Appointments <- Appointments %>%
  mutate(ReservationYear = 2019)


# Combine ReservationYear, ReservationMonth, and ReservationDay to create ReservationDate
# Combine CreationYear, CreationMonth, and CreationDay to create CreationDate
Appointments <- Appointments %>%
  mutate(ReservationDate = make_date(year = ReservationYear, month = ReservationMonth, day = ReservationDay),
         CreationDate = make_date(year = CreationYear, month = CreationMonth, day = CreationDay))


# Calculate the difference in days between ReservationDate and CreationDate
# Convert the difftime object to numeric
Appointments <- Appointments %>%
  mutate(ReservationSpan = as.numeric(difftime(ReservationDate, CreationDate, units = "days")))

```




# Exploratory Analysis


```{r}
# View summary of data before converting ReservationSpan to numeric
summary(Appointments)

# Convert ReservationSpan to numeric
Appointments <- Appointments %>%
  mutate(ReservationSpan = as.numeric(ReservationSpan))

# View summary of data after converting ReservationSpan to numeric
summary(Appointments)

```



```{r}
# Calculate average time between appointment creation and reservation date
average_reservation_span <- mean(Appointments$ReservationSpan)

# Find maximum time between appointment creation and reservation date
max_reservation_span <- max(Appointments$ReservationSpan)

# Find minimum time between appointment creation and reservation date
min_reservation_span <- min(Appointments$ReservationSpan)

# Print results
cat("a. Average time between appointment creation and reservation date:", average_reservation_span, "days\n")
cat("b. Maximum time between appointment creation and reservation date:", max_reservation_span, "days\n")
cat("c. Minimum time between appointment creation and reservation date:", min_reservation_span, "days\n")


```

__a. What is the average time between appointment creation and the actual reservation date?__
Approximately 38.53 days.
__b. What was the max time?__
207 days.
__c. What was the min time?__
1 day.




```{r}

# Calculate correlation between ReservationSpan and Show
correlation <- cor(Appointments$ReservationSpan, Appointments$Show)
print(correlation)  # Print correlation coefficient

# Create bar chart for Show variable
ggplot(Appointments, aes(x = factor(Show))) +
  geom_bar() +
  labs(x = "Show", y = "Frequency", title = "Frequency of Show vs No Show")

```



__a. Are ReservationSpan and Show highly correlated?__  
The correlation coefficient between `ReservationSpan` and `Show` is 0.002693853. Since this value is very close to zero, it indicates a very weak correlation between these two variables. Therefore, `ReservationSpan` and `Show` are not highly correlated.




__b. Based on 0 being “no show” and 1 being “show”, did most people make their appointments?__  
Yes, most people made their appointments.

