---
output:
  word_document: default
  html_document: default
---
# Module 3 - Assignment 1

## Pierce, Michaela

### Data Visualization


In this assignment, I will be using datasets containing candy rankings and production. These datasets can be found on the Canvas course website.

        
         

```{r}

library(tidyverse) 
candy_data <- read_csv("candy_data.csv") 
candy_production <- read_csv("candy_production.csv")

```


---

        
         

#### __Visualization with Scatterplots (geom_point)__

```{r}
# Create scatterplot (geom_point)
ggplot(candy_data, aes(x = sugarpercent, label= competitorname, y = pricepercent)) +
  geom_point()+
  geom_smooth(method = "lm") + # adds a fitted line 
  
labs(title = "Sugar by Price Scatterplot", # plot title 
  x = "Sugar content (percentile)", # x axis label 
  y = "Price (percentile)" # y axis label 
) +

geom_text(check_overlap = T,# automatically reduce overlap (deletes some labels) 
  vjust = "bottom", # adjust the vertical orientation 
  nudge_y = 0.01, # move the text up a bit so it doesn't touch the points 
  angle = 30,# tilt the text 30 degrees 
  size = 2 # make the text smaller (to reduce overlap more) 
) # and then add labels to the points 

```







The candy with the most sugar and the lowest price is ReeseÕs stuffed with pieces with a price of 0.651 % and a sugar content of 0.988 %. The most expensive candy with the highest sugar content is 100 Grand with a price of 0.86 % and a sugar content of 0.732 %.


Candy with the most sugar: ReeseÕs stuffed with pieces  
Candy with the least sugar: One dime  
Candy with the lowest price: Tootsie Roll Midgies  
Candy with the highest price: Nik L Nip   
    
       

  
---   

      
      
I wanted to get the exact values so it would be accurate. All the sources and code I used are below.


- Learned how to use cat to print, which converts and prints all arguments as a 
character string.      
<https://www.geeksforgeeks.org/printing-output-of-an-r-program/>
   
   
- Learned how to print a new line in a string. Found this is similar/same as printing new lines in Python.        
<https://www.geeksforgeeks.org/r-program-to-print-a-new-line-in-string/>     

- For Max / Refresher/Guide          
<https://www.geeksforgeeks.org/return-the-index-of-the-first-maximum-value-of-a-numeric-vector-in-r-programming-which-max-function/>

- For Min / Refresher/ Guide           
<https://www.geeksforgeeks.org/return-the-index-of-the-first-minimum-value-of-a-numeric-vector-in-r-programming-which-min-function/>



```{r}

# Find the index of the candy with the highest price
max_price_index <- which.max(candy_data$pricepercent)

# Find the index of the candy with the highest sugar content
max_sugar_index <- which.max(candy_data$sugarpercent)

# Find the index of the candy with the highest price among candies with the highest sugar content
max_price_highest_sugar_index <- which.max(candy_data$pricepercent[candy_data$sugarpercent == candy_data$sugarpercent[max_sugar_index]])

# Print the candy with the highest price and the most sugar
cat("Candy with the highest price and most sugar:", candy_data$competitorname[max_price_highest_sugar_index], "\n")
cat("Price:", candy_data$pricepercent[max_price_highest_sugar_index], "%\n")
cat("Sugar content:", candy_data$sugarpercent[max_price_highest_sugar_index], "%\n")

# Print the candy with the most sugar and lowest price
cat("Candy with the most sugar and lowest price:", candy_data$competitorname[max_sugar_index], "\n")
cat("Price:", candy_data$pricepercent[max_sugar_index], "%\n")
cat("Sugar content:", candy_data$sugarpercent[max_sugar_index], "%\n")


# Find the index of the candy with the most sugar
most_sugar_index <- which.max(candy_data$sugarpercent)
# Find the index of the candy with the least sugar
least_sugar_index <- which.min(candy_data$sugarpercent)
# Find the index of the candy with the lowest price
lowest_price_index <- which.min(candy_data$pricepercent)
# Find the index of the candy with the highest price
highest_price_index <- which.max(candy_data$pricepercent)

# Print the candies with the most and least sugar
cat("Candy with the most sugar:", candy_data$competitorname[most_sugar_index], "\n")
cat("Candy with the least sugar:", candy_data$competitorname[least_sugar_index], "\n")

# Print the candies with the lowest and highest price
cat("Candy with the lowest price:", candy_data$competitorname[lowest_price_index], "\n")
cat("Candy with the highest price:", candy_data$competitorname[highest_price_index], "\n")



```





   
   
---


      
          
#### __Line Chart of Candy Production__ 


I am using the candy production dataset that will display a specific 
date and how production during that month is compared to 2012.    

```{r}
# Create line chart with fitted line
ggplot(candy_production, aes(x = observation_date, y = IPG3113N)) +
  geom_line() +
  geom_smooth() +
  labs(title = "Monthly Candy Production",
       x = "Year",
       y = "As percent of 2012 production")
```



---


       
#### __Bar Chart of Ingredients__
   

```{r}

# Create bar chart for chocolate variable
ggplot(candy_data, aes(x = chocolate)) +
  geom_bar() 

# select out the columns that have the features of the candy (chocolate, caramel, etc.) 
candyFeatures <- candy_data %>% select(2:10) 
# make sure that these are booleans (logical) 
candyFeatures[] <- lapply(candyFeatures, as.logical)


# Create bar chart for candyFeatures dataset
ggplot(candyFeatures, aes(x = chocolate)) +
  geom_bar()

```













