---
output:
  word_document: default
  html_document: default
---
# Module 2 - Assignment 2 
## Pierce, Michaela
### Introduction to R Markdown



This is my first R Markdown document. It will go over some previous topics covered in class such as atomic vectors, data frames, and data types.  




The code I will be entering will contain a basic plot showing sales over the past 6 months.



```{r Plot of Sales by Month}

library(tidyverse)
sales <- c(50, 150, 200, 100, 300, 250) 
month <- c(1, 2, 3, 4, 5, 6)
qplot(month, sales, geom = "point", xlab = "Month", ylab= "Sales")


```



What month had the largest sales? Month 5  

What was the amount? 300  










```{r Working with Data Frames}



month <- c("Jan", "Feb", "Mar", "Apr", "May", "Jun", "Jul", "Aug", "Sep", "Oct", "Nov", "Dec")
sales <- c(150.25, 258.54, 268.55, 122.52, 987, 458.82, 667.23, 845.54, 586.78, 888.58, 756.12, 456.84)
Yearly_sales <- data.frame(month, sales)
     
library(ggplot2)
qplot(data = Yearly_sales, y = Yearly_sales$month, x = Yearly_sales$sales,main = "Monthly Sales", xlab = "Sales",ylab = "Month")
     


```




  
Which month had the most sales? May = 987  
Which month had the least sales? Apr = 122.52


